---
title: 'Visualize and interpret kinetics data'
author: "Melanie (1028-02098-001-001)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    depth: 3
    theme: sandstone
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{1028-02098-001-001}
- \fancyfoot[LE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options:
  chunk_output_type: console
---

# Introduction

The goal of this document is to visualize the kinetics simulations produced earlier and to summarize outcomes of accessibility and binding to interpret different modes of cooperativity.

# Computational Setup

```{r, warning=F, message=F}
#Standard packages
library(rtracklayer) ; library(GenomicRanges); library(magrittr) ; library(Biostrings)
library(ggplot2) ; library(reshape2); library(plyranges); library(Rsamtools); library(parallel)
library(dplyr); library(data.table); library(patchwork); library(readr); library(testit)

#KNITR Options
setwd("/n/projects/mw2098/publications/2024_weilert_acc/code/2_analysis/")
figure_filepath<-"figures/19_visualize_kinetics"
options(knitr.figure_dir=figure_filepath, java.parameters = "- Xmx6g")

#Lab sources
source("scripts/r/granges_common.r")
source("scripts/r/metapeak_common.r")
source("scripts/r/knitr_common.r")
source("scripts/r/caching.r")
source("scripts/r/metapeak_functions.r")

#Specific sources
library(BSgenome.Mmusculus.UCSC.mm10)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(ggseqlogo)
library(ggrastr)
library(basicdrm)
source("scripts/r/motif_functions.r")

#Pre-existing variables
bsgenome<-BSgenome.Mmusculus.UCSC.mm10
txdb<-TxDb.Mmusculus.UCSC.mm10.knownGene
kinetics.path<-'tsv/kinetics/kinetics_8state_scenarios.tsv.gz'

conc_levels.df<-readr::read_tsv('tsv/kinetics/conc_by_seq.tsv.gz') %>% 
  dplyr::mutate(conc_levels = 1:nrow(.),
                log10_kon = log10(kon))
conc_levels_n<-nrow(conc_levels.df)
```

# Plot S/S

In the event of plausible Sox2/Sox2 motif binding and pioneering, what cooperativity emerges from the accessibility (as represented by p(open) at the two states)?

```{r}
#Set limits
marginal_limits <- c(0, .85)
joint_limits <- c(0, .85)
```
We want to explain some features. For this, we will classify certain components.

```{r}
s_vs_s.df<-readr::read_tsv('tsv/kinetics/kinetics_8state_scenarios_SvsS.tsv.gz') 

s_vs_s_acc.plot<-ggplot(s_vs_s.df, aes(x = acc_marg, y = acc_joint))+
  geom_abline(slope = 1, intercept = 0, linetype = 'dashed')+
  ggrastr::geom_point_rast(size = .1)+
  # ggrastr::geom_point_rast(aes(color = a1Bvals + a1Avals), size = .1)+
  scale_x_continuous(limits = marginal_limits, name = 'Marginal p(open)')+
  scale_y_continuous(limits = joint_limits, name = 'Joint p(open)')+
  
  ggtitle('S vs S')+
  theme_classic() + theme(legend.position = 'none')
s_vs_s_acc.plot
```

# Plot S/OS

In the event of plausible Sox2/Oct4-Sox2 motif binding and pioneering, what cooperativity emerges from the accessibility (as represented by p(open) at the two states)?

```{r}
s_vs_os.df<-readr::read_tsv('tsv/kinetics/kinetics_8state_scenarios_SvsOS.tsv.gz') %>%
  dplyr::mutate(a1Avals_bin = cut(a1Avals, breaks = c(0, 3, 7, 10), include.lowest = T))

s_vs_os_acc.plot<-ggplot(s_vs_os.df, aes(x = acc_marg, y = acc_joint))+
  geom_abline(slope = 1, intercept = 0, linetype = 'dashed')+
  ggrastr::geom_point_rast(size = .1)+
  # ggrastr::geom_point_rast(aes(color = a1Avals_bin), size = .1)+
  scale_x_continuous(limits = marginal_limits, name = 'Marginal p(open)')+
  scale_y_continuous(limits = joint_limits, name = 'Joint p(open)')+
  # scale_color_manual(values = viridis(n=3))+
  ggtitle('S vs OS')+
  theme_classic() + theme(legend.position = 'none')
s_vs_os_acc.plot
```

Consolidate and output.

```{r}
plot<-s_vs_s_acc.plot + s_vs_os_acc.plot + plot_layout(nrow = 1, widths = c(1, 1))
ggsave(paste0(figure_filepath, '/joint_vs_marg_S_vs_OS.png'), height = 4, width = 8)
ggsave(paste0(figure_filepath, '/joint_vs_marg_S_vs_OS.pdf'), height = 4, width = 8)
```

# Plot concentration effects

Here, we will select a series of parameters over changing concentrations such that we can show (1) accessibility is higher when two TFs pioneer than when one does (this is superadditive) and (2) cooperativity (relative) peaks at different concentrations for different affinities.

## S/TFB when S changes and TF B does not

We will select a scenario with a less tight nucleosome and moderate pioneering. 

Plot Sox2 effects (A) when concentration of Sox2 (A) is changing and ONLY affinities of Sox2 (A) is changing in the presence of another pioneering TF (B) with unchanging affinities and concentration.

```{r}
conc.df<-Sys.glob('tsv/kinetics/os_sameclosing=True_*_a1=4.00-6.00_a2f=3.25-5.50_d=0.01-0.01_betam=5.tsv.gz') %>%
  lapply(., function(x){

    #Rows are trials, comma separators are the concentrations
    coop.df<-readr::read_tsv(x) %>% tidyr::separate(col = 'acc_coopvals', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
      dplyr::mutate(trial = 1:nrow(.)) %>%
      tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'acc_coopvals') %>% dplyr::select(acc_coopvals, conc_levels, trial)%>%
      dplyr::mutate(acc_coopvals = as.numeric(acc_coopvals))
    A_only.df<-readr::read_tsv(x) %>% tidyr::separate(col = 'acc_Aonly', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
      dplyr::mutate(trial = 1:nrow(.)) %>%
      tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'acc_Aonly') %>% dplyr::select(acc_Aonly, conc_levels, trial) %>%
      dplyr::mutate(acc_Aonly = as.numeric(acc_Aonly))
    B_only.df<-readr::read_tsv(x) %>% tidyr::separate(col = 'acc_Bonly', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
      dplyr::mutate(trial = 1:nrow(.)) %>%
      tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'acc_Bonly') %>% dplyr::select(acc_Bonly, conc_levels, trial) %>%
      dplyr::mutate(acc_Bonly = as.numeric(acc_Bonly))
    AB.df<-readr::read_tsv(x) %>% tidyr::separate(col = 'acc_AB', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
      dplyr::mutate(trial = 1:nrow(.)) %>%
      tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'acc_AB') %>% dplyr::select(acc_AB, conc_levels, trial) %>%
      dplyr::mutate(acc_AB = as.numeric(acc_AB))
    
    
    bindAonly.df<-readr::read_tsv(x) %>% tidyr::separate(col = 'bind_Aonly', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
      dplyr::mutate(trial = 1:nrow(.)) %>%
      tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'bind_Aonly') %>% dplyr::select(bind_Aonly, conc_levels, trial) %>%
      dplyr::mutate(bind_Aonly = as.numeric(bind_Aonly))
        
    bindBonly.df<-readr::read_tsv(x) %>% tidyr::separate(col = 'bind_Bonly', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
      dplyr::mutate(trial = 1:nrow(.)) %>%
      tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'bind_Bonly') %>% dplyr::select(bind_Bonly, conc_levels, trial) %>%
      dplyr::mutate(bind_Bonly = as.numeric(bind_Bonly))
            
    bindA_AB.df<-readr::read_tsv(x) %>% tidyr::separate(col = 'bindA_ABonly', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
      dplyr::mutate(trial = 1:nrow(.)) %>%
      tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'bindA_ABonly') %>% dplyr::select(bindA_ABonly, conc_levels, trial) %>%
      dplyr::mutate(bindA_ABonly = as.numeric(bindA_ABonly))
    
    #Assign parameters in case people are interested
     parameter_order<-c('kon', 'konB', 'koff1nA','koff1nB','gA','gB','betanA', 'betanB', 'betamA', 'betamB','h0','h1', 'kopen0','kclose0', 'a1nA', 'a1nB', 'a1nAB', 'a1mA', 'a1mB', 'a1mAnB', 'a1nAmB', 'a1mAmB', 'a2nA', 'a2nB', 'a2nAB', 'a2mA', 'a2mB', 'a2mAnB', 'a2nAmB', 'a2mAmB')
    parameters.df<-readr::read_tsv(x) %>% tidyr::separate(col = 'parset', sep = ',', into = parameter_order) %>%
      dplyr::mutate(trial = 1:nrow(.)) %>% dplyr::select(trial, parameter_order) %>% dplyr::select(-kon)
    
    df<-coop.df %>% 
      left_join(A_only.df) %>% 
      left_join(B_only.df) %>% 
      left_join(AB.df) %>%
      
      left_join(bindAonly.df) %>% 
      left_join(bindBonly.df) %>% 
      left_join(bindA_AB.df) %>%
      
      dplyr::left_join(parameters.df) %>% 
      dplyr::mutate(g_class = basename(x) %>% stringr::str_split(., '_') %>% unlist() %>% .[3]) %>%
      dplyr::mutate(conc_levels = as.integer(conc_levels)) %>%
      dplyr::left_join(conc_levels.df)
    
    return(df)
}) %>% rbindlist()


colors<-RColorBrewer::brewer.pal(n=4, name="Blues")
colors<-c('#bdc9e1','#74a9cf','#2b8cbe','#045a8d')
#Visualize features
single.plot<-ggplot(conc.df, aes(x = log10_kon, y = acc_Aonly, color = g_class, group = interaction(trial, g_class)))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on) of A')+
  scale_y_continuous(name = 'Acc. A only', limits = c(0, 1))+
  scale_color_manual(values = rev(colors), name = 'Affinity group')+
  theme_classic() + theme(legend.position = 'bottom')
print(max(conc.df$acc_Aonly))

singleB.plot<-ggplot(conc.df, aes(x = log10_kon, y = acc_Bonly, color = g_class, group = interaction(trial, g_class)))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on) of A')+
  scale_y_continuous(name = 'Acc. B only', limits = c(0, 1))+
  scale_color_manual(values = rev(colors), name = 'Affinity group')+
  theme_classic() + theme(legend.position = 'bottom')
print(max(conc.df$acc_Bonly))


double.plot<-ggplot(conc.df, aes(x = log10_kon, y = acc_AB, color = g_class, group = interaction(trial, g_class)))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on)')+
  scale_y_continuous(name = 'Acc. AB together', limits = c(0, 1))+
  scale_color_manual(values = rev(colors), name = 'Affinity group')+
  theme_classic() + theme(legend.position = 'bottom')
print(max(conc.df$acc_AB))

coop.plot<-ggplot(conc.df, aes(x = log10_kon, y = acc_coopvals, color = g_class, group = interaction(trial, g_class)))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on)')+
  scale_y_continuous(name = 'Acc. Coop', limits = c(0, 2))+
  scale_color_manual(values = rev(colors), name = 'Affinity group')+
  theme_classic() + theme(legend.position = 'bottom')


conc.plots<-single.plot + singleB.plot + double.plot + coop.plot + plot_layout(nrow = 1) + plot_annotation(title = 'S/TFB where S affinity/conc change', subtitle = 'a1=4.00-6.00_a2f=3.25-5.50_d=0.01-0.01_betam=5')
conc.plots
ggsave(paste0(figure_filepath, '/conc_S_vs_TFB_open_region_mid_pioneering.png'), conc.plots, height = 4, width = 12)
ggsave(paste0(figure_filepath, '/conc_S_vs_TFB_open_region_mid_pioneering.pdf'), conc.plots, height = 4, width = 12)
```

Plot binding occupancy over changing concentrations

```{r}
#Visualize features
single.plot<-ggplot(conc.df, aes(x = log10_kon, y = bind_Aonly, color = g_class, group = interaction(trial, g_class)))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on) of A')+
  scale_y_continuous(name = 'Bind. A only', limits = c(0, 1))+
  scale_color_manual(values = rev(colors), name = 'Affinity group')+
  theme_classic() + theme(legend.position = 'bottom')

singleB.plot<-ggplot(conc.df, aes(x = log10_kon, y = bind_Bonly, color = g_class, group = interaction(trial, g_class)))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on) of A')+
  scale_y_continuous(name = 'Bind. B only', limits = c(0, 1))+
  scale_color_manual(values = rev(colors), name = 'Affinity group')+
  theme_classic() + theme(legend.position = 'bottom')


double.plot<-ggplot(conc.df, aes(x = log10_kon, y = bindA_ABonly, color = g_class, group = interaction(trial, g_class)))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on)')+
  scale_y_continuous(name = 'Bind. AB together', limits = c(0, 1))+
  scale_color_manual(values = rev(colors), name = 'Affinity group')+
  theme_classic() + theme(legend.position = 'bottom')


conc.plots<-single.plot + singleB.plot + double.plot + plot_layout(nrow = 1) + plot_annotation(title = 'S/TFB where S affinity/conc change', 
                                                                                               subtitle = 'a1=4.00-6.00_a2f=3.25-5.50_d=0.01-0.01_betam=5')
conc.plots
ggsave(paste0(figure_filepath, '/conc_S_vs_TFB_open_region_mid_pioneering_binding_occupancy.png'), conc.plots, height = 4, width = 12)
ggsave(paste0(figure_filepath, '/conc_S_vs_TFB_open_region_mid_pioneering_binding_occupancy.pdf'), conc.plots, height = 4, width = 12)

```

Plot the curve derivatives.

```{r}
conc_w_dx.df<-conc.df %>%
  dplyr::mutate(accAonly_dx = c(NA, diff(acc_Aonly)),
                acc_AB_dx = c(NA, diff(acc_AB))) %>%
  tidyr::pivot_longer(cols = c(accAonly_dx, acc_AB_dx), names_to = 'measurement', values_to = 'dx') %>%
  as.data.frame

#Visualize features
dx.plot<-ggplot(conc_w_dx.df, aes(x = log10_kon, y = dx, color = g_class, group = interaction(trial, g_class)))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on) of A')+
  scale_y_continuous(name = 'Dx of curves', limits = c(0, 0.015))+
  facet_grid(. ~ measurement)+
  scale_color_manual(values = rev(colors), name = 'Affinity group')+
  ggtitle('S/TFB where S affinity/conc change', subtitle = 'a1=4.00-6.00_a2f=3.25-5.50_d=0.01-0.01_betam=5')+
  theme_classic() + theme(legend.position = 'bottom')
dx.plot
ggsave(paste0(figure_filepath, '/conc_S_vs_TFB_open_region_mid_pioneering_curve_derivatives.png'), dx.plot, height = 4, width = 8)
ggsave(paste0(figure_filepath, '/conc_S_vs_TFB_open_region_mid_pioneering_curve_derivatives.pdf'), dx.plot, height = 4, width = 8)
```

#### Hill coefficients

Calculate Hill coefficients for each of these features. Better documentation comes from another package (https://search.r-project.org/CRAN/refmans/braidrm/html/findBestHill.html). 

```{r, eval = F}
#Calculate median scores across trials
conc_summary.df<-conc.df %>%
  dplyr::group_by(g_class, kon) %>%
  dplyr::summarize(med_acc_Aonly = median(acc_Aonly),
                   med_acc_Bonly = median(acc_Bonly),
                   med_acc_AB = median(acc_AB))

hill_fit_model<-'m3plc'

#Calculate Hill coefficient where max is the plateau ceiling
max_is_ceiling.df<-lapply(conc_summary.df$g_class %>% unique, function(x){
  c.df<-conc_summary.df %>% dplyr::filter(g_class==x)

  fit<-basicdrm::findBestHillModel(formula = c.df$kon, data = c.df$med_acc_Aonly,
                                   defaults = c(0, max(c.df$med_acc_Aonly)))
  A.df<-fit$allfits[[hill_fit_model]]$coefficients %>% t()  %>% as.data.frame() %>%
    dplyr::mutate(gA = x,
                  state = 'A_only')


  fit<-basicdrm::findBestHillModel(formula = c.df$kon, data = c.df$med_acc_Bonly,,
                                   defaults = c(0, max(c.df$med_acc_Bonly)))
  B.df<-fit$allfits[[hill_fit_model]]$coefficients %>% t()  %>% as.data.frame() %>%
    dplyr::mutate(gA = x,
                  state = 'B_only')

  fit<-basicdrm::findBestHillModel(formula = c.df$kon, data = c.df$med_acc_AB,,
                                   defaults = c(0, max(c.df$med_acc_AB)))
  AB.df<-fit$allfits[[hill_fit_model]]$coefficients %>% t()  %>% as.data.frame() %>%
    dplyr::mutate(gA = x,
                  state = 'AB')

  return(rbind(A.df, B.df, AB.df))
}) %>% rbindlist()
print(max_is_ceiling.df)

#Calculate Hill coefficient where p(open)=1 is the plateau ceiling
p100_is_ceiling.df<-lapply(conc_summary.df$g_class %>% unique, function(x){
  c.df<-conc_summary.df %>% dplyr::filter(g_class==x)

  fit<-basicdrm::findBestHillModel(formula = c.df$kon, data = c.df$med_acc_Aonly,
                                   defaults = c(0, 1))
  A.df<-fit$allfits[[hill_fit_model]]$coefficients %>% t()  %>% as.data.frame() %>%
    dplyr::mutate(gA = x,
                  state = 'A_only')


  fit<-basicdrm::findBestHillModel(formula = c.df$kon, data = c.df$med_acc_Bonly,,
                                   defaults = c(0, 1))
  B.df<-fit$allfits[[hill_fit_model]]$coefficients %>% t()  %>% as.data.frame() %>%
    dplyr::mutate(gA = x,
                  state = 'B_only')

  fit<-basicdrm::findBestHillModel(formula = c.df$kon, data = c.df$med_acc_AB,,
                                   defaults = c(0,1))
  AB.df<-fit$allfits[[hill_fit_model]]$coefficients %>% t()  %>% as.data.frame() %>%
    dplyr::mutate(gA = x,
                  state = 'AB')

  return(rbind(A.df, B.df, AB.df))
}) %>% rbindlist()
print(p100_is_ceiling.df)
```

Confirms that Hill coefficient of this curve is 1 or 10 (non log space) regardless of how you calculate it. Alone vs cooperative doesn't change that.

## S/TFB when both S and TFB can change affinities and concentrations

Generate plots where we change gA/gB in a two TF conformation (high/high, high/low, low/high, low/low) across concentrations where (A change, B change, both change) to show that motif affinity and concentration play in higher order fashions due to cooperativity. 


```{r}

scenarios.df<-readr::read_tsv('tsv/kinetics/S_vs_TFB_changing_variables.tsv')
scenarios.df$konA %>% summary()
# colors3<-RColorBrewer::brewer.pal(n=3, name="Blues")
colors3<-c('#bdc9e1','#74a9cf','#045a8d')


acc.plot<-scenarios.df %>% 
  dplyr::filter((conc_scenario=='A_change_B_static')|(konA==konB)) %>%
  dplyr::filter(affinity_scenario != 'gA_low_gB_high') %>%
  ggplot(., aes(x = log10(konA), y = acc_AB, color = affinity_scenario))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on) of A')+
  scale_y_continuous(name = 'Acc', limits = c(0, 1))+
  scale_color_manual(values = rev(colors3), name = 'Affinity group')+
  facet_grid(. ~ conc_scenario)+
  theme_classic() + theme(legend.position = 'right')

coop.plot<-scenarios.df %>% 
  dplyr::filter((conc_scenario=='A_change_B_static')|(konA==konB)) %>%
  dplyr::filter(affinity_scenario != 'gA_low_gB_high') %>%
  ggplot(., aes(x = log10(konA), y = acc_coopvals, color = affinity_scenario))+
  geom_line(size = 1) + 
  scale_x_continuous(name = 'log10(k_on) of A')+
  scale_y_continuous(name = 'Coop', limits = c(-.5, 2))+
  scale_color_manual(values = rev(colors3), name = 'Affinity group')+
  facet_grid(. ~ conc_scenario)+
  theme_classic() + theme(legend.position = 'right')


changing_modes<-acc.plot + coop.plot + plot_layout(nrow = 1) + plot_annotation(title = 'S/TFB where both affinity/conc change', subtitle = 'a1=4.00-6.00_a2f=3.25-5.50_d=0.01-0.01_betam=5')
changing_modes
ggsave(paste0(figure_filepath, '/conc_S_vs_TFB_open_region_both_change.png'), changing_modes, height = 4, width = 12)
ggsave(paste0(figure_filepath, '/conc_S_vs_TFB_open_region_both_change.pdf'), changing_modes, height = 4, width = 12)
```

# Run parameter controls (Sox2 alone)

## Non-equilibrium

Show that g, o, c, betam, and d have differential effects on concentration curves according to our models. 

```{r}
controls.df<-readr::read_tsv('tsv/kinetics/kinetics_8state_parameter_controls_for_Aonly_sox2.tsv') 
control_key.list<-list(affinity = 'gA', opening = 'a1mA', closing = 'a2mA', d = 'kclose0', betam = 'betamA')

controls_formatted.plot<-lapply(controls.df$control %>% unique, function(x){
  print(x)
  
  #Rows are trials, comma separators are the concentrations
  A_only.df<-controls.df %>% dplyr::filter(control==x) %>% dplyr::mutate(parameter_range = 1:nrow(.)) %>% tidyr::separate(col = 'acc_Aonly', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
    tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'acc_Aonly') %>% 
    dplyr::select(acc_Aonly, conc_levels, parameter_range) %>%
    dplyr::mutate(acc_Aonly = as.numeric(acc_Aonly))
  
  B_only.df<-controls.df %>% dplyr::filter(control==x) %>% dplyr::mutate(parameter_range = 1:nrow(.)) %>% tidyr::separate(col = 'acc_Bonly', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
    tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'acc_Bonly') %>% 
    dplyr::select(acc_Bonly, conc_levels, parameter_range) %>%
    dplyr::mutate(acc_Bonly = as.numeric(acc_Bonly))
  
  #Assign parameters in case people are interested
  parameter_order<-c('kon', 'konB', 'koff1nA','koff1nB','gA','gB','betanA', 'betanB', 'betamA', 'betamB','h0','h1', 'kopen0','kclose0', 'a1nA', 'a1nB', 'a1nAB', 'a1mA', 'a1mB', 'a1mAnB', 'a1nAmB', 'a1mAmB', 'a2nA', 'a2nB', 'a2nAB', 'a2mA', 'a2mB', 'a2mAnB', 'a2nAmB', 'a2mAmB')
  parameters.df<-controls.df %>% dplyr::filter(control==x) %>% dplyr::mutate(parameter_range = 1:nrow(.)) %>% tidyr::separate(col = 'parset', sep = ',', into = parameter_order) %>% dplyr::select(parameter_order, parameter_range) %>% dplyr::select(-kon) %>% as.data.frame()
  
  df<-A_only.df %>% 
    dplyr::left_join(B_only.df) %>%
    dplyr::left_join(parameters.df) %>% 
    dplyr::mutate(parameter_of_interest = as.numeric(!!sym(control_key.list[[x]]))) %>%
    dplyr::select(acc_Aonly, acc_Bonly, conc_levels, parameter_range, parameter_of_interest) %>% 
    dplyr::mutate(conc_levels = as.integer(conc_levels)) %>%
    dplyr::left_join(conc_levels.df)
  
  #Custom process values to project them fairly
  if(x=='d'){
    df$parameter_of_interest_final<-0.05/df$parameter_of_interest
    x = 'k_close'
  } 
  # else if (x=='closing') {
  #   df$parameter_of_interest_final<-1/df$parameter_of_interest
  #   x = '1/close'
  # } 
  else {
    df$parameter_of_interest_final<-df$parameter_of_interest
  }
  
  # colors<-c('#bdc9e1','#74a9cf','#2b8cbe','#045a8d', '#08306b')
  #Visualize features
  single.plot<-ggplot(df, aes(x = log10_kon, y = acc_Aonly, color = parameter_of_interest_final, group = parameter_of_interest_final))+
    geom_line(aes(y = acc_Bonly, group = parameter_of_interest_final), color = 'red', size = 1) + 
    geom_line(size = 1) + 
    geom_hline(yintercept = 1, linetype = 'dashed')+
    scale_x_continuous(name = 'log10(k_on) of A')+
    scale_y_continuous(name = 'Acc. A only', limits = c(0, 1))+
    scale_color_gradient(high = '#3f007d', low = '#bcbddc', name = x)+
    # scale_color_manual(values = rev(colors), name = x)+
    # scale_color_manual(values = rev(RColorBrewer::brewer.pal(df$parameter_of_interest %>% unique %>% length, "Blues")), name = x)+
    ggtitle('Sox2 on Sox2 motif', subtitle = x)+
    theme_classic() + theme(legend.position = 'bottom')
  
  if(x %in% c('k_close', 'affinity')){
    single.plot<-single.plot + scale_color_gradient(high = '#bcbddc', low = '#3f007d', name = x, trans = "log")
  }
  if(x %in% c('closing')){
    single.plot<-single.plot + scale_color_gradient(low = '#bcbddc', high = '#3f007d', name = x, trans = "log")
  }
  
  return(single.plot)
  
}) %>% wrap_plots(nrow = 1)

controls_formatted.plot
ggsave(paste0(figure_filepath, '/conc_S_parameter_controls.png'), controls_formatted.plot, height = 4, width = 20)
ggsave(paste0(figure_filepath, '/conc_S_parameter_controls.pdf'), controls_formatted.plot, height = 4, width = 20)
```

## Equilibrium

Show that g, betam, and d have differential effects on concentration curves according to our models. 

```{r}
controls.df<-readr::read_tsv('tsv/kinetics/kinetics_4state_parameter_controls_for_equilibrium_sox2.tsv') 
control_key.list<-list(affinity = 'gA', d = 'kclose0', betam = 'betamA')

controls_formatted.plot<-lapply(controls.df$control %>% unique, function(x){
  print(x)
  
  #Rows are trials, comma separators are the concentrations
  A_only.df<-controls.df %>% dplyr::filter(control==x) %>% dplyr::mutate(parameter_range = 1:nrow(.)) %>% tidyr::separate(col = 'acc_A', sep = ',', into = c(1:conc_levels_n) %>% as.character) %>%
    tidyr::pivot_longer(cols = c(1:conc_levels_n) %>% as.character, names_to = 'conc_levels', values_to = 'acc_A') %>% 
    dplyr::select(acc_A, conc_levels, parameter_range) %>%
    dplyr::mutate(acc_A = as.numeric(acc_A))
  
  #Assign parameters in case people are interested
  parameter_order<-c('kon', 'koff1nA','gA','betanA', 'betamA','kopen0','kclose0', 'a1nA', 'a1mA','a2nA', 'a2mA')
  parameters.df<-controls.df %>% dplyr::filter(control==x) %>% dplyr::mutate(parameter_range = 1:nrow(.)) %>% tidyr::separate(col = 'parset', sep = ',', into = parameter_order) %>% dplyr::select(parameter_order, parameter_range) %>% dplyr::select(-kon) %>% as.data.frame()
  
  df<-A_only.df %>% 
    dplyr::left_join(parameters.df) %>% 
    dplyr::mutate(parameter_of_interest = as.numeric(!!sym(control_key.list[[x]]))) %>%
    dplyr::select(acc_A, conc_levels, parameter_range, parameter_of_interest) %>% 
    dplyr::mutate(conc_levels = as.integer(conc_levels)) %>%
    dplyr::left_join(conc_levels.df)
  
  
    #Custom process values to project them fairly
  if(x=='d'){
    df$parameter_of_interest_final<-0.05/df$parameter_of_interest
    x = 'k_close'
  } else {
    df$parameter_of_interest_final<-df$parameter_of_interest
  }
  
  #Visualize features
  single.plot<-ggplot(df, aes(x = log10_kon, y = acc_A, color = parameter_of_interest_final, group = parameter_of_interest_final))+
    geom_line(size = 1) + 
    geom_hline(yintercept = 1, linetype = 'dashed')+
    scale_x_continuous(name = 'log10(k_on) of A')+
    scale_y_continuous(name = 'Acc. A only', limits = c(0, 1))+
    scale_color_gradient(high = '#3f007d', low = '#bcbddc', name = x)+
    # scale_color_manual(values = rev(colors), name = x)+
    # scale_color_manual(values = rev(RColorBrewer::brewer.pal(df$parameter_of_interest %>% unique %>% length, "Blues")), name = x)+
    ggtitle('Sox2 on Sox2 motif', subtitle = x)+
    theme_classic() + theme(legend.position = 'bottom')
  
  if(x %in% c('k_close', 'affinity')){
    single.plot<-single.plot + scale_color_gradient(high = '#bcbddc', low = '#3f007d', name = x, trans = "log")
  }
  if(x %in% c('closing')){
    single.plot<-single.plot + scale_color_gradient(low = '#bcbddc', high = '#3f007d', name = x, trans = "log")
  }
  return(single.plot)
  
}) %>% wrap_plots(nrow = 1)

controls_formatted.plot
ggsave(paste0(figure_filepath, '/conc_S_parameter_controls_equilibrium.png'), controls_formatted.plot, height = 4, width = 20)
ggsave(paste0(figure_filepath, '/conc_S_parameter_controls_equilibrium.pdf'), controls_formatted.plot, height = 4, width = 20)
```

# Session Information

For the purposes of reproducibility, the R/Bioconductor session information is printed below:

```{r sessioninfo}
sessionInfo()
```












