---
title: 'Neuronal differentiation analysis'
author: "Melanie (1028-02098-001-001)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    depth: 3
    theme: sandstone
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{1028-02098-001-001}
- \fancyfoot[LE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options:
  chunk_output_type: console
---

# Introduction

The goal of this analysis is to:

1. Map day0 motifs (ESCs) from day12 motifs (neural progenitors)
2. See whether there is a change from cooperative interactions of Sox2 with Oct4-Sox2 (loss of Oct4) to a different motif candidate that is specific to neural progenitors.

# Computational Setup

```{r, warning=F, message=F}
#Standard packages
library(rtracklayer) ; library(GenomicRanges); library(magrittr) ; library(Biostrings)
library(ggplot2) ; library(reshape2); library(plyranges); library(Rsamtools); library(parallel)
library(dplyr); library(data.table); library(patchwork); library(readr); library(testit)

#KNITR Options
setwd("/n/projects/mw2098/publications/2024_weilert_acc/code/2_analysis/")
figure_filepath<-"figures/20_bunina_analysis"
options(knitr.figure_dir=figure_filepath, java.parameters = "- Xmx6g")

#Lab sources
source("scripts/r/granges_common.r")
source("scripts/r/metapeak_common.r")
source("scripts/r/knitr_common.r")
source("scripts/r/caching.r")
source("scripts/r/metapeak_functions.r")

#Specific sources
library(BSgenome.Mmusculus.UCSC.mm10)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(org.Mm.eg.db)
library(ggseqlogo)
library(DESeq2)
source("scripts/r/motif_functions.r")

#Pre-existing variables
threads <- 5
bsgenome<-BSgenome.Mmusculus.UCSC.mm10
txdb<-TxDb.Mmusculus.UCSC.mm10.knownGene

#model information
peaks.path<-'bed/bpreveal/bunina_atac_day0_across_combined_fold1_all.bed'
input_length<-2032
output_length<-1000

shap.bw.list<-list(`day0` = 'shap/bunina_atac_day0_across_combined_fold1_atac_counts.bw',
                   `day4` = 'shap/bunina_atac_day4_across_combined_fold1_atac_counts.bw',
                   `day8` = 'shap/bunina_atac_day8_across_combined_fold1_atac_counts.bw',
                   `day12` = 'shap/bunina_atac_day12_across_combined_fold1_atac_counts.bw')
```

# Curate motifs

From fold1 of day0 and day12, identify motifs that are commonly occuring.

```{r}
seqlet_ids.list<-list(
  fold1 = list(
    bunina_atac_day0 = data.frame(
      metacluster_name = 'pos_patterns', # Need this if you also have negative patterns
      pattern_name = c('pattern_0', 'pattern_1', 'pattern_2', 'pattern_3', 'pattern_4',
                       'pattern_5','pattern_6', 'pattern_7', 'pattern_8', 'pattern_9',
                       'pattern_10', 'pattern_11', 'pattern_12', 'pattern_13', 'pattern_14',
                       'pattern_15'), 
      motif = c('Sp1', 'Nfy', 'Klf4', 'Oct4-Sox2', 'Ctcf', 
                'Sox2','Nrf1', 'Creb', 'Oct4-Sox2','Nrf1',
                'Elk', 'Ctcf', 'Ctcf', 'Thap11', 'Tead',
                'Zbtb33'),
      oriented_strand = c('+', '-' , '+', '+', '-',
                          '+', '+', '+', '+', '+',
                          '-', '-', '-', '+', '+',
                          '+'),
      task = 'atac'),
    
    bunina_atac_day8 = data.frame(
      metacluster_name = 'pos_patterns', # Need this if you also have negative patterns
      pattern_name = c('pattern_0', 'pattern_1', 'pattern_2', 'pattern_3', 'pattern_4',
                       'pattern_5','pattern_6', 'pattern_7', 'pattern_8', 'pattern_9',
                       'pattern_10', 'pattern_11', 'pattern_12', 'pattern_13', 'pattern_14',
                       'pattern_15', 'pattern_16'), 
      motif = c('Sp1', 'Nfy', 'Nrf1', 'Crcf', 'Sox2', 
                'Elk','Creb', 'Thap11', 'Rfx', 'Pbx2/Meis',
                'Zbtb33', 'TGAGCTA', 'Elk-TGAGCTA', 'Elk-Elk', 'Rfx',
                'Rfx', 'Rfx'),
      oriented_strand = c('+', '+' , '+', '-', '+',
                          '-', '+', '+', '+', '+',
                          '+', '-', '-', '+', '+',
                          '+', '+'),
      task = 'atac'),
    
    bunina_atac_day12 = data.frame(
      metacluster_name = 'pos_patterns', # Need this if you also have negative patterns
      pattern_name = c('pattern_0', 'pattern_1', 'pattern_2', 'pattern_3', 'pattern_4',
                       'pattern_5','pattern_6', 'pattern_7', 'pattern_8', 'pattern_9',
                       'pattern_10', 'pattern_11', 'pattern_12'), 
      motif = c('Sp1', 'Nfy', 'Nrf1', 'Creb', 'Elk', 
                'Thap11', 'Rfx','Ctcf', 'Lhx', 'Cbx',
                'Sox-like','Ctcf', 'Ctcf'),
      oriented_strand = c('+', '-' , '+', '+', '-', 
                          '+', '+', '+', '+', '+', 
                          '+', '+', '+'),
      task = 'atac')
  )
)

seqlet_ids.df<-lapply(seqlet_ids.list, function(x){
  x %>% rbindlist(idcol = 'model')
}) %>% rbindlist(idcol = 'fold')
readr::write_tsv(seqlet_ids.df, 'tsv/mapped_motifs/bunina_pattern_ids.tsv')
```

## Import hits

Using fold1, import all hits from the key motifs of interest.

```{r}
hits.list<-lapply(names(seqlet_ids.list$fold1), function(model){
  df<-lapply(Sys.glob(paste0('modiscolite/', model, '*_fold1_*counts/hits.tsv')),
                     function(x) readr::read_tsv(x) %>%
                       dplyr::mutate(model = model,
                                     fold = 'fold1')) %>% 
    rbindlist() %>% as.data.frame %>% 
    dplyr::left_join(., seqlet_ids.df) %>%
    dplyr::filter(!is.na(motif))
  print(table(df$motif))
  return(df)
})
names(hits.list)<-names(seqlet_ids.list$fold1)
```

## Curate motifs redundancy

Remove the following redundancies:

+ Motifs that are peak-redundant will be mapped across peak "indexes" that are actually located across overlapping genomic regions. 
+ Motifs that are palindromic will be mapped on the positive and negative strand with almost identical ranges. 
  + Here, we will remove redundantly mapped motifs and select the unique motif maps that have the highest contribution.
+ Motifs that have sub-components such as Sox2 mapped over an Oct4-Sox2 motif. We will assign Oct4-Sox2 motifs as the prior.
+ Motifs that have sub-components such as Nanog mapped over an Oct4-Sox2/Sox2 motif. We will assign Oct4-Sox2/Sox2 motifs as the prior.

In order to simultaneously manage peak-redundancy and palindromic motifs, we will have to assign generalized "peak islands" intead of `peak_index` as our grouped region sets.

```{r}
hits_dedup.list<-lapply(names(hits.list), function(model){
  #Orient strand to match desired aligned convention
  gr<-makeGRangesFromDataFrame(hits.list[[model]], keep.extra.columns = T, starts.in.df.are.0based = T) %>%
    plyranges::mutate(strand = ifelse(oriented_strand == '-', 
                                      ifelse(strand == '-', '+', '-'),
                                      strand))
  
  #Import peaks and reduce them to a deduplicated state to get general peak regions.
  peaks.gr<-rtracklayer::import(peaks.path) %>% 
    GenomicRanges::resize(., input_length, fix = 'center') %>% 
    GenomicRanges::reduce(., ignore.strand = T) %>% 
    plyranges::mutate(region_id_tmp = 1:length(.))
  
  #Find overlaps between motifs and peaks
  ov<-findOverlaps(gr, peaks.gr, ignore.strand = T)
  
  #Assign temporary region ID for deduplication
  gr$region_id_tmp<-NA
  gr$region_id_tmp[ov@from]<-peaks.gr$region_id_tmp[ov@to]

  #Filter any motif that doesn't fit into peaks. This means that the motif is mapped on the edges of the general, temporary region_id
  gr<-gr %>% plyranges::filter(!is.na(gr$region_id_tmp))
  
  #Deduplicate any motifs for palindromic mappings
  dedup.gr<-lapply(unique(gr$motif), function(y){
    print(y)
    motifs.gr<-gr %>% plyranges::filter(motif==y)
    motifs_dedup.gr<-remove_palindromic_motifs_from_bpnet_instances(dfi = motifs.gr, 
                                                                    pattern_to_filter = y, 
                                                                    chrom_name = 'seqnames', start_name = 'start', end_name = 'end', 
                                                                    value_name = 'contrib_match', motif_name = 'motif',
                                                                    region_name = 'region_id_tmp')
    print(paste0(length(motifs_dedup.gr) , ' kept from ', length(motifs.gr)))
    return(motifs_dedup.gr)
  }) %>% as('GRangesList') %>% unlist(.)

  final_motifs.gr<-dedup.gr %>%
    plyranges::select(-c(oriented_strand, fold))
  
  return(final_motifs.gr)
})
names(hits_dedup.list)<-names(hits.list)
```

## Double check motif overlaps

Ensure that there isn't an overrepresentation of unexpected overlapping motif pairs. We expect some overlaps, but checking to make sure that our curations are consistent. 

```{r}
lapply(hits_dedup.list, function(gr){
  reduced_motifs.gr<-gr %>% GenomicRanges::reduce(., ignore.strand = T, min.gapwidth = 0)
  gr$motif_id_tmp<-findOverlaps(gr, reduced_motifs.gr, ignore.strand = T)@to
  motifs_overlapped_ids<-table(gr$motif_id_tmp)[table(gr$motif_id_tmp)>1]
  
  gr %>%
    as.data.frame %>%
    dplyr::filter(motif_id_tmp %in% names(motifs_overlapped_ids)) %>%
    dplyr::group_by(motif_id_tmp) %>%
    dplyr::summarize(overlapping_motif_name = paste0(sort(motif), collapse = ':')) %>%
    .$overlapping_motif_name %>%
    table(.) %>% 
    sort(decreasing = T) %>%
    head(n=30)
})
```

## Remove redundant motif instances

```{r}
hits_no_redundant.list<-lapply(names(hits_dedup.list), function(model){
  gr<-hits_dedup.list[[model]]
  no_redundant.gr<-remove_redundant_motifs_from_bpnet_instances(gr,chrom_name = 'seqnames', start_name = 'start', end_name = 'end',
                                                                value_name = 'contrib_magnitude',
                                                                region_name = 'region_id_tmp')
  
  return(no_redundant.gr)
})
names(hits_no_redundant.list)<-names(hits.list)
```

Compare changed numbers in motifs

```{r}
lapply(hits_dedup.list, function(x) x$motif %>% table)
lapply(hits_no_redundant.list, function(x) x$motif %>% table)
```

We see a loss in the motifs we would expect to had they been redundant.

## Calculate islands

Because we took motifs based on their relative contribution and not the featured `peak_index` column, we need a new grouping annotation to define the motifs when we are conducting perturbations and other grouping systems. We will call this index `region_id`.

```{r}
motifs_w_annotations.gr.list<-lapply(names(hits_no_redundant.list), function(model){
  message(model)
  
  gr<-hits_no_redundant.list[[model]]
  
  islands.gr<-resize(gr, 150, 'center') %>%
    GenomicRanges::reduce(ignore.strand = T)
  # rtracklayer::export(islands.gr, 'bed/mapped_motifs/islands_raw.bed')
  
  data.frame(x = width(islands.gr)) %>%
    ggplot(., aes(x = x))+
    geom_density()
  
  #Filter islands that are >1000 bp
  islands.gr <- islands.gr %>% plyranges::filter(width <= input_length)  %>%
    plyranges::mutate(region_id = 0:(length(.)-1)) #Make region_id 0-based in order to be compatible with BPNet contribution .h5 file indexing properties.
  ov<-findOverlaps(gr, islands.gr, ignore.strand = T)
  islands_resized_to_output.gr <- islands.gr %>% GenomicRanges::resize(., width = output_length, fix = 'center')
  
  #Nice! No regions get filtered out this time. This is great.
  
  #Export regions in 0-based coordinates
  rtracklayer::export(islands.gr %>% 
                        plyranges::mutate(name = region_id), 
                      paste0('bed/mapped_motifs/all_', model, '_islands_curated_0based.bed'), format = 'BED')
  rtracklayer::export(islands.gr %>% 
                        GenomicRanges::resize(., width = output_length, fix = 'center') %>% 
                        plyranges::mutate(name = region_id), 
                      paste0('bed/mapped_motifs/all_', model, '_islands_curated_0based_sized_to_output.bed'), format = 'BED')
  rtracklayer::export(islands.gr %>% 
                        GenomicRanges::resize(., width = input_length, fix = 'center') %>% 
                        plyranges::mutate(name = region_id), 
                      paste0('bed/mapped_motifs/all_', model, '_islands_curated_0based_sized_to_input.bed'), format = 'BED')
  
  #Assign region_id to these motifs
  gr_w_regions.gr<-gr
  gr_w_regions.gr$region_id<-NA
  gr_w_regions.gr$region_id[ov@from]<-islands_resized_to_output.gr$region_id[ov@to]
  gr_w_regions.gr$region_start_1based<-NA
  gr_w_regions.gr$region_start_1based[ov@from]<-start(islands_resized_to_output.gr)[ov@to]
  gr_w_regions.gr$region_end_1based<-NA
  gr_w_regions.gr$region_end_1based[ov@from]<-end(islands_resized_to_output.gr)[ov@to]
  
  #Filter motifs that aren't in a sufficiently small island.
  gr_w_ic.gr <- gr_w_regions.gr %>% plyranges::filter(!is.na(region_id)) %>%
    plyranges::mutate(motif_id = 0:(length(.)-1), #0-based indexing later might come in handy
                      seq = getSeq(bsgenome, ., as.character = T))
  
  #1-based
  gr_w_ic.df<-gr_w_ic.gr %>% as.data.frame()
  readr::write_tsv(gr_w_ic.df, paste0('tsv/mapped_motifs/all_', model, '_instances_curated_1based.tsv.gz'))
  
  #0-based
  rtracklayer::export(gr_w_ic.gr %>% plyranges::mutate(name = paste0(motif, '_', motif_id, '_', region_id)), 
                      paste0('bed/mapped_motifs/all_', model, '_instances_curated_0based.bed'))
  return(gr_w_ic.gr)
  
})
names(motifs_w_annotations.gr.list)<-names(hits.list)
```

# Plot SHAP scores of different concentrations across mapped motifs

```{r}
motifs_of_interest.gr<-c(
  motifs_w_annotations.gr.list$bunina_atac_day0 %>% plyranges::filter(motif=='Oct4-Sox2', pattern_name== 'pattern_3'),
  motifs_w_annotations.gr.list$bunina_atac_day0 %>% plyranges::filter(motif=='Ctcf', pattern_name== 'pattern_4'),
  motifs_w_annotations.gr.list$bunina_atac_day12 %>% plyranges::filter(motif=='Rfx')
)


cwms.plot.list<-lapply(motifs_of_interest.gr$motif %>% unique, function(x){
  message(x)
  logos.plot.list<-mclapply(names(shap.bw.list), function(y){
    message(y)
      gr<-motifs_of_interest.gr %>% plyranges::filter(motif==x)
      seqs<-gr %>% getSeq(bsgenome, ., as.character = T)
      contrib.mat<-standard_metapeak_matrix(regions.gr = gr,
                                            sample.cov = shap.bw.list[[y]],
                                            keep_region_coordinates = T)
      seqs_1he.list<-lapply(seqs, function(z) one_hot_encode_DNA(z))
      contribs.list<-lapply(1:dim(contrib.mat)[1], function(z){
        seqs_1he.list[[z]] * matrix(rep(contrib.mat[z,], 4), nrow = 4, byrow = T)
      })
      cwm<-Reduce("+", contribs.list) / length(contribs.list)
      
      cwm.plot<-ggseqlogo(cwm, method = 'custom', seq_type='dna')+
        scale_x_continuous(name = 'Motif position (bp)')+
        scale_y_continuous(name = 'Contrib', limits = c(-.005, .04))+
        ggtitle(paste0(x, '_', y)) + 
        theme(axis.line.y = element_line(), axis.ticks.y = element_line(),
              axis.title = element_blank(), axis.text.x = element_blank())
    return(cwm.plot)
  }, mc.cores = 2)
  
  logos.plot<-wrap_plots(logos.plot.list) + plot_layout(ncol = 1)
  ggsave(paste0(figure_filepath, '/cwm_across_mapped_', x, '.png'), logos.plot, height = 9, width = 3)
  ggsave(paste0(figure_filepath, '/cwm_across_mapped_', x, '.pdf'), logos.plot, height = 9, width = 3)
  
  return(logos.plot)
}) 

all_logos.plot<-wrap_plots(cwms.plot.list) + plot_layout(nrow = 1)
all_logos.plot
ggsave(paste0(figure_filepath, '/cwm_across_mapped_all.png'), all_logos.plot, height = 11, width = 12)
ggsave(paste0(figure_filepath, '/cwm_across_mapped_all.pdf'), all_logos.plot, height = 11, width = 12)
```

# Find candidates of Sox2 cooperativity

```{r}
#Find regions that have Rfx and Oct4-Sox2 nearby
subsetByOverlaps(motifs_w_annotations.gr.list$bunina_atac_day0 %>% plyranges::filter(motif %in% c('Oct4-Sox2')),
                 motifs_w_annotations.gr.list$bunina_atac_day8 %>% plyranges::filter(motif=='Sox2')) %>%
  GenomicRanges::resize(., 400, 'center') %>%
  subsetByOverlaps(.,
                   motifs_w_annotations.gr.list$bunina_atac_day8 %>% plyranges::filter(motif=='Pbx2/Meis')) %>%
  plyranges::arrange(desc(contrib_magnitude)) %>% as.data.frame


#chr11:71,996,116-71,996,273 #immediate cooperativity thats super nice bw sox2 and rfx
#chr13:84,059,688-84,059,837 great candidate showing perfect cooperativity handover
#chr2:117,938,915-117,939,042 secondary candidate 
#chr10:99,538,846-99,538,942 good secondary candidate
```

# GM17750 intronic enhancer showcase

## Showcase the contribution tracks

```{r}
curated_enhancer_boundaries.gr<- GRanges('chr13', IRanges(84059730,84059820))
```

Generate contribution scores of mapped motifs.

```{r}
acc_contrib.plot<-lapply(c(0, 8), function(x){
  
  curated_enhancer.df<-data.frame(
    acc_contrib = standard_metapeak_matrix(regions.gr =curated_enhancer_boundaries.gr, sample.cov = paste0('shap/bunina_atac_day', x, '_across_combined_fold1_atac_counts.bw'), 
                                           keep_region_coordinates = T) %>% t(),
    seq = getSeq(bsgenome, curated_enhancer_boundaries.gr, as.character = T) %>% stringr::str_split_1(pattern = ''),
    position = start(curated_enhancer_boundaries.gr):end(curated_enhancer_boundaries.gr)
  )
  
  # Generate sequence logo
  acc_contrib.plot<-curated_enhancer.df%>%
    tidyr::pivot_wider(names_from = seq, values_from = acc_contrib) %>% 
    dplyr::select(A, C, G, T) %>% as.matrix() %>% t() %>%
    ggseqlogo(., method='custom', seq_type='dna') + 
    scale_y_continuous(name = 'Acc. contrib.')+
    scale_x_discrete(name = 'Genomic position (chr13, bp)')+
    ggtitle(paste0('day ', x))
}) %>% wrap_plots(ncol = 1)

acc_contrib.plot
ggsave(paste0(figure_filepath, '/GM17750_contrib.png'), acc_contrib.plot, height = 3, width = 10)
ggsave(paste0(figure_filepath, '/GM17750_contrib.pdf'), acc_contrib.plot, height = 3, width = 10)
```

## Showcase cooperativity between timepoints

Arrange motifs

```{r}
candidate_island.df<-curated_enhancer_boundaries.gr %>% plyranges::mutate(region_id = 0, comb_max = 2, comb_min = 1) %>% GenomicRanges::resize(., width = 2033, 'center') %>% as.data.frame()%>%
  dplyr::rename(chrom = seqnames) 
candidate_motifs.df<-c(subsetByOverlaps(motifs_w_annotations.gr.list$bunina_atac_day0, curated_enhancer_boundaries.gr, ignore.strand = T),
  subsetByOverlaps(motifs_w_annotations.gr.list$bunina_atac_day8, curated_enhancer_boundaries.gr, ignore.strand = T)) %>%
  as.data.frame() %>%
  plyranges::mutate(region_id = 0, 
                    start = start, 
                    end = end+1,
                    motif_window_start = start - candidate_island.df$start,
                    motif_window_end = end - candidate_island.df$start)  %>%
  dplyr::rename(chrom = seqnames) #all motifs are on negative strand
readr::write_tsv(candidate_motifs.df, 
                 'tsv/genomic/motifs/bunina_showcase_instances_formatted_for_genomic_perturbations_0based.tsv.gz')
readr::write_tsv(candidate_island.df,
                 'tsv/genomic/motifs/bunina_showcase_islands_formatted_for_genomic_perturbations_0based.tsv.gz')
```

Calculate cooperativity

```{bash, eval = F}
source /home/mw2098/.bashrc
conda deactivate
conda activate bpreveal_404
cd /n/projects/mw2098/publications/2024_weilert_acc/code/2_analysis

/home/mw2098/anaconda3/envs/bpreveal_404/bin/python scripts/py/tools/generate_genomic_perturbs.py \
--motifs_df_path tsv/genomic/motifs/bunina_showcase_instances_formatted_for_genomic_perturbations_0based.tsv.gz \
--motif_groups_df_path tsv/genomic/motifs/bunina_showcase_islands_formatted_for_genomic_perturbations_0based.tsv.gz \
--model_dir models/bunina_atac_day0_across_combined_fold1_combined.model / \
--output_prefix tsv/genomic/motifs/bunina_showcase_instances_day0 \
--tasks_in_model_order atac \
--region_index_column region_id \
--motif_window_start_column motif_window_start \
--motif_window_end_column motif_window_end \
--motif_name_column motif \
--motif_chrom_column chrom \
--fasta_file ../0_setup/fa/mm10.fa \
--input_seqlen 2032 --output_seqlen 1000 \
--comb_max comb_max --comb_min comb_min \
--nodes 1 --trials 16 \
--use_whole_window 

/home/mw2098/anaconda3/envs/bpreveal_404/bin/python scripts/py/tools/generate_genomic_perturbs.py \
--motifs_df_path tsv/genomic/motifs/bunina_showcase_instances_formatted_for_genomic_perturbations_0based.tsv.gz \
--motif_groups_df_path tsv/genomic/motifs/bunina_showcase_islands_formatted_for_genomic_perturbations_0based.tsv.gz \
--model_dir models/bunina_atac_day8_across_combined_fold1_combined.model / \
--output_prefix tsv/genomic/motifs/bunina_showcase_instances_day8 \
--tasks_in_model_order atac \
--region_index_column region_id \
--motif_window_start_column motif_window_start \
--motif_window_end_column motif_window_end \
--motif_name_column motif \
--motif_chrom_column chrom \
--fasta_file ../0_setup/fa/mm10.fa \
--input_seqlen 2032 --output_seqlen 1000 \
--comb_max comb_max --comb_min comb_min \
--nodes 4 --trials 16 \
--use_whole_window 
```

Quantify and print cooperativity

```{r}
perturb.df<-readr::read_tsv('tsv/genomic/motifs/bunina_showcase_instances_day8_chr13.tsv.gz')

data.frame(AB = perturb.df %>% dplyr::filter(mut=='Reference') %>% .$`atac/pred_sum`,
           A = perturb.df %>% dplyr::filter(mut=='Sox2-1') %>% .$`atac/pred_sum`,
           B = perturb.df %>% dplyr::filter(mut=='Pbx2/Meis-1') %>% .$`atac/pred_sum`,
           null = perturb.df %>% dplyr::filter(mut=='Pbx2/Meis-1_Sox2-1') %>% .$`atac/pred_sum`) %>%
  dplyr::mutate(coop = (AB-null)/(A + B - 2*null),
                log_coop = log(coop))

```

# Run DESeq2

Across this differentiation time course, run DESeq on accessibility and RNA-seq levels.

## Define reference regions

Import regions that you are doing to calculate differential enrichment across. For this, we will take a union set of the different timepoints and resize them to a systematic level.

```{r}
regions.gr<-Sys.glob(paste0('bed/mapped_motifs/all_bunina_*_islands_curated_0based.bed')) %>% lapply(., function(x) rtracklayer::import(x)) %>%
  as('GRangesList') %>%
  unlist() %>%
  GenomicRanges::reduce(ignore.strand = T) %>%
  GenomicRanges::resize(., input_length, 'center') 
```

## Collect DESeq2 normalized accessibility across regions

```{r}
#Collect data
atac.bw.vec<-Sys.glob('../1_processing/bw/Bunina_mesc_atac_day*_*_cutsites.bw')
atac.df<-mclapply(atac.bw.vec, function(x){
  regionSums(regions.gr, 
             x)
}, mc.cores = 2) %>% as.data.frame()

#Format organization
atac_labels.df<-data.frame(filename = atac.bw.vec) %>%
  dplyr::mutate(state = basename(filename) %>% gsub('Bunina_mesc_atac_', '', .) %>% gsub('_cutsites.bw', '', .)) %>%
  tidyr::separate(state, into = c('timepoint','replicate'), remove = F)

#Assign names to counts table
names(atac.df)<-atac_labels.df$state

#Run DESeq2
dds <- DESeqDataSetFromMatrix(countData = atac.df,
                              colData = atac_labels.df,
                              design = ~ timepoint)
dds <- DESeq(dds)

#Return enrichment results
log2FoldChange_atac.df<-lapply(paste0('day', seq(4, 12, 4)), function(x){
  res <- results(dds, contrast=c("timepoint",x, "day0"))
  res$log2FoldChange
}) %>% as.data.frame()
colnames(log2FoldChange_atac.df)<-paste0('log2fc_atac_day', seq(4, 12, 4), 'h_vs_day0')
head(log2FoldChange_atac.df)

#Return whether the region is significantly up- or down-regulated
padj_atac.df<-lapply(paste0('day', seq(4, 12, 4)), function(x){
  res <- results(dds, contrast=c("timepoint",x, "day0"))
  res$padj
}) %>% as.data.frame()
colnames(padj_atac.df)<-paste0('padj_atac_day', seq(4, 12, 4), 'h_vs_day0')
head(padj_atac.df)

regions_w_deseq.df<-cbind(regions.gr %>% as.data.frame, log2FoldChange_atac.df, padj_atac.df)
```

## Collect DESeq2 normalized transcription across enhancers

```{r}
#Collect data
erna.bw.vec<-Sys.glob('../1_processing/bw/Bunina_mesc_rna_day*_*.bw')
erna.df<-mclapply(erna.bw.vec, function(x){
  regionSums(regions.gr,
             x)
}, mc.cores = 2) %>% as.data.frame()

#Format organization
erna_labels.df<-data.frame(filename = erna.bw.vec) %>%
  dplyr::mutate(state = basename(filename) %>% gsub('Bunina_mesc_rna_', '', .) %>% gsub('.bw', '', .)) %>%
  tidyr::separate(state, into = c('timepoint','replicate'), remove = F)
names(erna.df)<-erna_labels.df$state

#Run DESeq2
dds <- DESeqDataSetFromMatrix(countData = erna.df,
                              colData = erna_labels.df,
                              design = ~ timepoint)
dds <- DESeq(dds)

#Return enrichment results
log2FoldChange_erna.df<-lapply(paste0('day', seq(4, 12, 4)), function(x){
  res <- results(dds, contrast=c("timepoint",x, "day0"))
  res$log2FoldChange
}) %>% as.data.frame()
colnames(log2FoldChange_erna.df)<-paste0('log2fc_rna_day', seq(4, 12, 4), 'h_vs_day0')

#Return whether the region is significantly up- or down-regulated
padj_erna.df<-lapply(paste0('day', seq(4, 12, 4)), function(x){
  res <- results(dds, contrast=c("timepoint",x, "day0"))
  res$padj
}) %>% as.data.frame()
colnames(padj_erna.df)<-paste0('padj_rna_day', seq(4, 12, 4), 'h_vs_day0')
head(padj_erna.df)

regions_w_deseq_both.df<-cbind(regions_w_deseq.df %>% as.data.frame, log2FoldChange_erna.df, padj_erna.df)
head(regions_w_deseq_both.df)
readr::write_tsv(regions_w_deseq_both.df, 'tsv/mapped_motifs/all_bunina_islands_curated_1based_w_deseq2_data.tsv.gz')
```

## Collect DESeq2 normalized transcription across genes

Given published TT-seq data from Bunina et al, assess pairwise differential expression across annotated regions. This will be a differentiation time course. 

```{r}
#Collect data
rnaseq.counts.path<-Sys.glob('../1_processing/bam/Bunina_mesc_rna_day*ReadsPerGene.out.tab')
rnaseq_conditions.vec<-rnaseq.counts.path %>% basename %>% gsub('Bunina_mesc_rna_', '', .) %>% gsub('ReadsPerGene.out.tab', '', .)
rnaseq_annotations.gr<-rtracklayer::import('../../public/databases/Ens/mm10.Ens_98.gtf')

#Import counts tables
rnaseq_names.vec<-readr::read_tsv(rnaseq.counts.path[1], F) %>% .[[1]]
rnaseq.df<-mclapply(rnaseq.counts.path, function(x){
  df<-readr::read_tsv(x, F)
  colnames(df)<-c('gene_name','reads','fwd','rev')
  df$reads
}, mc.cores = 2) %>% as.data.frame()
colnames(rnaseq.df)<-rnaseq_conditions.vec
rnaseq.df$gene_id<-rnaseq_names.vec
rnaseq.df<-rnaseq.df %>% dplyr::filter(!grepl('N_', gene_id))

#Filter out genes that have no reads across them
genes_to_keep.vec<-rnaseq.df %>% tidyr::pivot_longer(cols = rnaseq_conditions.vec, names_to = 'condition', values_to = 'counts') %>%
  dplyr::group_by(gene_id) %>% dplyr::summarize(total_counts = sum(counts)) %>%
  dplyr::filter(total_counts>0) %>%
  .$gene_id %>% unique()

rnaseq.df<-rnaseq.df %>% dplyr::filter(gene_id %in% genes_to_keep.vec)
  
#Attach annotations
rnaseq_w_ann.df<-rnaseq.df %>%
  dplyr::left_join(., rnaseq_annotations.gr %>% as.data.frame %>% dplyr::select(gene_id, gene_name, gene_source)) %>%
  unique()
testit::assert(nrow(rnaseq_w_ann.df)==nrow(rnaseq.df))

#Format organization
rnaseq_labels.df<-data.frame(filename = rnaseq.counts.path) %>%
  dplyr::mutate(state = basename(filename) %>% basename %>% gsub('Bunina_mesc_rna_', '', .) %>% gsub('ReadsPerGene.out.tab', '', .)) %>%
  tidyr::separate(state, into = c('timepoint','replicate'), remove = F)

#Run DESeq2
dds <- DESeqDataSetFromMatrix(countData = rnaseq.df %>% dplyr::select(-gene_id),
                              colData = rnaseq_labels.df,
                              design = ~ timepoint)
dds <- DESeq(dds)

#Return enrichment results
log2FoldChange_rnaseq.df<-lapply(paste0('day', seq(4, 12, 4)), function(x){
  res <- results(dds, contrast=c("timepoint",x, "day0"))
  res$log2FoldChange
}) %>% as.data.frame()
colnames(log2FoldChange_rnaseq.df)<-paste0('log2fc_rna_gene_day', seq(4, 12, 4), 'h_vs_day0')

#Return significance results
signif_rnaseq.df<-lapply(paste0('day', seq(4, 12, 4)), function(x){
  res <- results(dds, contrast=c("timepoint",x, "day0"))
  res$padj
}) %>% as.data.frame()
colnames(signif_rnaseq.df)<-paste0('padj_rna_gene_day', seq(4, 12, 4), 'h_vs_day0')
```

Connect differential RNA-seq values with annotations. Then connect gene locations with the nearest motif groups.

```{r}
rnaseq_w_deseq.df<-cbind(rnaseq_w_ann.df, log2FoldChange_rnaseq.df) %>%
  dplyr::left_join(., rnaseq_annotations.gr %>% as.data.frame %>% dplyr::filter(type=='gene') %>% dplyr::select(gene_id, gene_name, gene_source, seqnames, start, end, strand))
readr::write_tsv(rnaseq_w_deseq.df, 'tsv/mapped_motifs/all_bunina_genes_curated_1based_w_deseq2_data.tsv.gz')

#Collect significance
rnaseq_w_signif.df<-cbind(rnaseq_w_ann.df, signif_rnaseq.df, log2FoldChange_rnaseq.df) %>%
  dplyr::left_join(., rnaseq_annotations.gr %>% as.data.frame %>% dplyr::filter(type=='gene') %>% dplyr::select(gene_id, gene_name, gene_source, seqnames, start, end, strand))
```

## Read in differential data

```{r}
rnaseq_w_deseq.df<-readr::read_tsv('tsv/mapped_motifs/all_bunina_genes_curated_1based_w_deseq2_data.tsv.gz')
regions_w_deseq_both.df<-readr::read_tsv('tsv/mapped_motifs/all_bunina_islands_curated_1based_w_deseq2_data.tsv.gz')
```

# Observed mESC-specific fold-change over time 

Given OS motifs derived from mESCs as an anchor, how does gene expression change with the same arrangements of grouped/isolated? 

```{r}
motifs_at_day0.gr<-motifs_w_annotations.gr.list$bunina_atac_day0 %>% 
  # plyranges::filter(motif %in% c('Klf4', 'Sox2', 'Oct4-Sox2')) %>%
  as.data.frame %>%
  dplyr::group_by(motif) %>%
  dplyr::mutate(seq_match_quantile = ecdf(seq_match)(seq_match),
                seq_match_quantile_bin = cut(seq_match_quantile, breaks = c(0, .33, .67, 1), include.lowest = T, labels = paste0('smq_', 1:3)),
                region_id = as.integer(region_id)) %>%
  dplyr::group_by(region_id) %>%
  dplyr::mutate(island_count = dplyr::n(), 
                island_content = paste(unique(sort(motif)), table(sort(motif)), sep = ':', collapse = '_'),
                island_content_ordered = paste(motif, collapse = '_'),
                island_content_unique = paste(unique(sort(motif)), collapse = '_')) %>%
  dplyr::ungroup() %>%
  makeGRangesFromDataFrame(keep.extra.columns = T, starts.in.df.are.0based = T, 
                           seqnames.field = 'seqnames', start.field = 'start', end.field = 'end', 
                           strand.field = 'strand')

subsetByOverlaps(motifs_at_day0.gr %>% plyranges::filter(motif=='Sox2'), motifs_at_day0.gr %>% plyranges::filter(motif=='Oct4-Sox2'), ignore.strand = T)
```

Filter motifs based on the rationale that the enhancers wont be re-engineered to be neural progenitors. So we will take regions that don't have neuronal motifs emerge.

```{r}
# specific_neuronal_motifs.gr<-motifs_w_annotations.gr.list$bunina_atac_day12 %>% plyranges::filter(motif %in% c('Cbx', 'Lhx', 'Rfx'))
# 
# #Attach region_id
regions_at_day0.gr<-rtracklayer::import('bed/mapped_motifs/all_bunina_atac_day0_islands_curated_0based_sized_to_input.bed')
# regions_at_day0.gr<-regions_at_day0.gr %>%
#   plyranges::filter(!overlapsAny(., specific_neuronal_motifs.gr, ignore.strand = T))
```

About half the regions are lost.


Assign region information.

```{r}
regions_w_deseq_both.gr<-regions_w_deseq_both.df %>% makeGRangesFromDataFrame(keep.extra.columns = T)

ov<-findOverlaps(regions_w_deseq_both.gr, regions_at_day0.gr, ignore.strand = T)
regions_w_deseq_both.gr$region_id<-NA
regions_w_deseq_both.gr$region_id[ov@from]<-regions_at_day0.gr$name[ov@to]

regions_w_deseq_both.df<-regions_w_deseq_both.gr %>% 
  plyranges::mutate(atac_obs = regionSums(., 'bw/Bunina_mesc_atac_day0_1_cutsites.bw'),
                    atac_obs_q = ecdf(atac_obs)(atac_obs),
                    atac_obs_q_bin = plyr::round_any(atac_obs_q, accuracy = .2, f = ceiling)) %>%
  as.data.frame %>% dplyr::filter(!is.na(region_id)) %>% dplyr::mutate(region_id = as.integer(region_id)) %>%
  dplyr::left_join(motifs_at_day0.gr %>% as.data.frame %>% dplyr::select(region_id, island_count, island_content, island_content_ordered, island_content_unique) %>% unique())
```

## Differential plots

### Observed accessibility data

```{r}
grouped.plot<-lapply(c('Oct4-Sox2', 'Sox2', 'Ctcf'), function(x){

  #Import motifs
  motif.df<-motifs_at_day0.gr %>% 
    as.data.frame %>% 
    dplyr::filter(motif==x) %>%
    dplyr::arrange(contrib_magnitude) %>%
    dplyr::mutate(row = 1:dplyr::n())
  
  #Measure effects of accessibility levels from DESeq2
  atac_comparisons<-colnames(log2FoldChange_atac.df)
  obs.df<-motif.df %>% dplyr::select(motif, region_id, row) %>%
    dplyr::left_join(., regions_w_deseq_both.df %>% dplyr::select(atac_comparisons, region_id)) %>%
    dplyr::filter(!is.na(log2fc_atac_day4h_vs_day0)) %>%
    dplyr::select(atac_comparisons, row) %>%
    tidyr::pivot_longer(cols = atac_comparisons,
                        names_to = 'timepoint', values_to = 'deseq_atac') %>%
    dplyr::mutate(timepoint = timepoint %>% factor(., levels = c(atac_comparisons)))

  #Collect islands belonging to each motif
  if(x=='Oct4-Sox2'){
    grouped.df<-obs.df %>% 
      dplyr::left_join(motif.df %>% dplyr::select(row, region_id, seq_match_quantile_bin, seq_match_quantile)) %>% 
      dplyr::left_join(regions_w_deseq_both.df %>% dplyr::select(region_id, island_count, island_content, atac_obs_q_bin)) %>%
      dplyr::mutate(island_state = ifelse(island_count==1, 'isolated', 'grouped')) %>%
      dplyr::filter(grepl('Oct4-Sox2:1', island_content))
  } else{
    grouped.df<-obs.df %>% 
      dplyr::left_join(motif.df %>% dplyr::select(row, region_id, seq_match_quantile_bin, seq_match_quantile)) %>% 
      dplyr::left_join(regions_w_deseq_both.df %>% dplyr::select(region_id, island_count, island_content, atac_obs_q_bin)) %>%
      dplyr::mutate(island_state = ifelse(island_count==1, 'isolated', 'grouped')) %>%
      dplyr::filter(!grepl('Oct4-Sox2', island_content))
  }
  
  motifs_w_grouped.df<-motif.df %>% dplyr::select(motif, region_id, row, seq_match_quantile_bin, seq_match_quantile) %>%
    dplyr::left_join(., regions_w_deseq_both.df %>% dplyr::select(atac_comparisons, region_id, atac_obs_q_bin, atac_obs_q, atac_obs)) %>%
    dplyr::left_join(grouped.df %>% dplyr::select(row, island_state) %>% unique()) %>% 
    dplyr::filter(atac_obs_q_bin>=.6) %>%
    dplyr::filter(!is.na(island_state))
  
  grouped.df<-grouped.df %>% dplyr::filter(atac_obs_q_bin>=.6)
  message(x, ': ', nrow(motif.df))
  print(table(motifs_w_grouped.df$island_state, motifs_w_grouped.df$seq_match_quantile_bin))

  #Plot distributions of isolated or grouped
  abs_distributions.plot<-motifs_w_grouped.df %>% as.data.frame %>%
    ggplot(.,aes(x = log(atac_obs), fill = island_state))+
    geom_density(alpha = .5)+
    scale_x_continuous(name = 'Abs accessibility')+# limits = c(5, 7.5))+
    scale_y_continuous(name = 'Density')+
    scale_fill_manual(values = c('#b2182b', '#2166ac'))+
    ggtitle('Abs accessibiltiy')+
    theme_classic()
  abs_distributions.plot
  ggsave(paste0(figure_filepath, '/motif_role_over_showcased_groups_', x, '_abs_accessibility_distributions.png'), abs_distributions.plot, height = 4, width = 6)
  ggsave(paste0(figure_filepath, '/motif_role_over_showcased_groups_', x, '_abs_accessibility_distributions.pdf'), abs_distributions.plot, height = 4, width = 6)
  # print(wilcox.test(motifs_w_grouped.df %>% dplyr::filter(island_state=='isolated') %>% .$atac_obs %>% log, grouped.df %>% dplyr::filter(island_state=='grouped') %>% .$atac_obs %>% log))
  
  #Plot absolute accessibility over categories
  my_comparisons<-list(c('smq_1','smq_3'))
  all_abs_acc.plot<-motifs_w_grouped.df %>% as.data.frame %>%
    ggplot(.,aes(x = seq_match_quantile_bin, y = log(atac_obs)))+
    geom_boxplot(outliers = F)+
    ggpubr::stat_compare_means(comparisons = my_comparisons, method = 'wilcox.test')+
    scale_y_continuous(name = 'Abs accessibility')+# limits = c(5, 7.5))+
    scale_x_discrete(name = 'Motif PWM group')+
    facet_grid(. ~ island_state)+
    ggtitle('Abs accessibiltiy')+
    theme_classic()
  all_abs_acc.plot
  ggsave(paste0(figure_filepath, '/motif_role_over_showcased_groups_', x, '_abs_accessibility.png'), all_abs_acc.plot, height = 4, width = 6)
  ggsave(paste0(figure_filepath, '/motif_role_over_showcased_groups_', x, '_abs_accessibility.pdf'), all_abs_acc.plot, height = 4, width = 6)

  #Plot  affinities over categories
  all_affinities.plot<-motifs_w_grouped.df %>% as.data.frame %>%
    
    #Remove all poor accessibility regions because we want to consider only open features. 
    #This cutoff was determined from absolute observed accessibility effects in later plots. See similar results despite changing cutoffs. 
    dplyr::filter(atac_obs_q_bin>=.6) %>%
    ggplot(.,aes(x = seq_match_quantile_bin, y = seq_match_quantile, fill = island_state))+
    geom_boxplot()+
    ggpubr::stat_compare_means(method = 'wilcox.test')+
    scale_y_continuous(name = 'PWM scores')+
    scale_x_discrete(name = 'Motif PWM group')+
    ggtitle('Affinities')+
    theme_classic()
  all_affinities.plot
  ggsave(paste0(figure_filepath, '/motif_role_over_showcased_groups_', x, '_affinities.png'), all_affinities.plot, height = 4, width = 6)
  ggsave(paste0(figure_filepath, '/motif_role_over_showcased_groups_', x, '_affinities.pdf'), all_affinities.plot, height = 4, width = 6)

  #Group and summarize and plot motifs by boolean (grouped/ungrouped) and again by affinity group
  grouped_deseq.plot<-grouped.df %>%
    dplyr::group_by(island_state, timepoint) %>%
    dplyr::summarize(med_deseq_atac = median(deseq_atac, na.rm = T)) %>%
    ggplot(., aes(x = timepoint, y = med_deseq_atac, color = island_state, group = interaction(island_state))) +
    # geom_point()+
    geom_line()+
    scale_x_discrete(name = 'Timepoint')+
    scale_y_continuous(name = 'Observed Log2FC (mut/0h)', limits = c(-2.5, .5))+
    theme_classic()
  
  affinity_deseq.plot<-grouped.df %>%
    dplyr::mutate(island_state = ifelse(island_count==1, 'isolated', 'grouped')) %>%
    dplyr::group_by(island_state, timepoint, seq_match_quantile_bin) %>%
    dplyr::summarize(med_deseq_atac = median(deseq_atac, na.rm = T)) %>%
    ggplot(., aes(x = timepoint, y = med_deseq_atac, linetype = island_state, color = seq_match_quantile_bin, group = interaction(island_state, seq_match_quantile_bin))) +
    # geom_point()+
    geom_line()+
    scale_x_discrete(name = 'Timepoint')+
    scale_y_continuous(name = 'Observed Log2FC (mut/0h)', limits = c(-2.5, .5))+
    theme_classic()
  
  #Save plots as a summary plot
  metaplot<-grouped_deseq.plot + affinity_deseq.plot
  ggsave(paste0(figure_filepath, '/motif_role_over_timepoints_', x, '_diff_accessibility_summary.png'), metaplot, height = 5, width = 10)
  ggsave(paste0(figure_filepath, '/motif_role_over_timepoints_', x, '_diff_accessibility_summary.pdf'), metaplot, height = 5, width = 10)

  return(NULL)
})
```

## Differential expression

### Observed differential gene expression

```{r}
#For all motifs, calculate the nearest gene occurrences
#When we subset, we will still take the only Oct4-Sox2 regulated regions that are nearby a gene ONE TIME within 100Kbp
motifs_via_occ.gr<-motifs_at_day0.gr %>% as.data.frame %>% dplyr::filter(motif %in% c('Oct4-Sox2', 'Sox2')) %>%
  dplyr::left_join(., regions_w_deseq_both.df %>% dplyr::select(region_id, island_content, island_count)) %>%
  makeGRangesFromDataFrame(keep.extra.columns = T, seqnames.field = 'seqnames', start.field = 'start', end.field = 'end')
rnaseq_w_deseq.gr<-rnaseq_w_deseq.df %>% makeGRangesFromDataFrame(keep.extra.columns = T)

motifs_via_occ.gr$gene_name<-rnaseq_w_deseq.gr$gene_name[nearest(motifs_via_occ.gr, rnaseq_w_deseq.gr, ignore.strand = T)]
motifs_via_occ.gr$nearest_gene_start<-rnaseq_w_deseq.df$start[nearest(motifs_via_occ.gr, rnaseq_w_deseq.gr, ignore.strand = T)]
motifs_via_occ.gr$nearest_gene_distance<-abs(motifs_via_occ.gr$nearest_gene_start - start(motifs_via_occ.gr))

motifs_via_occ.gr<-motifs_via_occ.gr %>% plyranges::filter(nearest_gene_distance<=100000)
gene_occurrence.df<-motifs_via_occ.gr %>% as.data.frame %>% 
  dplyr::select(region_id, gene_name) %>% unique %>% 
  dplyr::group_by(gene_name) %>% dplyr::summarize(gene_occurrence = dplyr::n())

rnaseq_comparisons<-paste0('log2fc_rna_gene_day', seq(4, 12, 4), 'h_vs_day0')
```

Plot magnitude of depletion of nearest target gene.

```{r}
nearest_gene_count_threshold<-20
grouped.plot<-lapply(c('Oct4-Sox2', 'Sox2'), function(x){
  
  motif.gr<-motifs_via_occ.gr %>% dplyr::filter(motif==x) %>%
    makeGRangesFromDataFrame(keep.extra.columns = T, seqnames.field = 'seqnames', start.field = 'start', end.field = 'end')
  

  nearest_thresholds<-c(5000, 50000, 100000)
  #Now, replot but with boundaries rather than simple minimums
  plot.list<-lapply(nearest_thresholds, function(thresh){
    message(thresh)
    lower_bound<-max(nearest_thresholds[nearest_thresholds<thresh])
    if(is.infinite(lower_bound)){lower_bound<-0}

    motif_near.gr<-motif.gr %>% plyranges::filter(nearest_gene_distance<=thresh,
                                                  nearest_gene_distance>=lower_bound)

    genes_to_consider.df<-rnaseq_w_deseq.gr %>% as.data.frame %>%
      dplyr::left_join(., gene_occurrence.df) %>%
      dplyr::mutate(gene_occurrence = ifelse(is.na(gene_occurrence), 0, gene_occurrence),
                    rnaseq_at_gene_0h = `day0_1` + `day0_2`) %>%
      dplyr::filter(gene_occurrence==1) %>%
      dplyr::left_join(motif_near.gr %>% as.data.frame %>% dplyr::select(gene_name, region_id, island_count, island_content) %>% unique,
                       ., by = 'gene_name') %>%
      dplyr::filter(!is.na(gene_occurrence)) %>%
      dplyr::mutate(island_state = ifelse(island_count==1, 'isolated', 'grouped'))

    #Collect islands belonging to each motif
    if(x=='Oct4-Sox2'){
      grouped.df<-genes_to_consider.df %>% dplyr::filter(grepl('Oct4-Sox2:1', island_content))
    } else{
      grouped.df<-genes_to_consider.df %>% dplyr::filter(!grepl('Oct4-Sox2', island_content))
    }
    readr::write_tsv(grouped.df, paste0('tsv/mapped_motifs/nearest_genes_bunina_curated_1based_wrt_', x, '_thresh_', as.integer(thresh), '.tsv.gz'))

    #Print numbers
    print(grouped.df %>% dplyr::select(region_id, island_state) %>% unique %>% 
            dplyr::group_by(island_state) %>% dplyr::summarize(counts = dplyr::n()))

    grouped.df<-grouped.df %>% tidyr::pivot_longer(cols = rnaseq_comparisons,
                                                   names_to = 'timepoint', values_to = 'deseq') %>%
      dplyr::mutate(timepoint = timepoint %>% factor(., levels = c(rnaseq_comparisons), labels = c(rnaseq_comparisons %>% gsub('deseq', '', .))))
    

    #Group and summarize and plot motifs by boolean (grouped/ungrouped) and again by affinity group
    group.plot<-grouped.df %>%
      dplyr::group_by(island_state, timepoint) %>%
      dplyr::summarize(med_deseq = median(deseq, na.rm = T),
                       freq = dplyr::n()) %>%
      dplyr::filter(freq>=nearest_gene_count_threshold) %>%
      ggplot(., aes(x = timepoint, y = med_deseq, color = island_state, group = interaction(island_state))) +
      geom_hline(yintercept = 0, linetype = 'dashed')+
      geom_point()+
      geom_line()+
      # facet_grid(gene_freq ~ .)+
      scale_x_discrete(name = 'Timepoint')+
      scale_y_continuous(name = 'Observed Log2FC (mut/0h)', limits = c(-.5, .5))+
      ggtitle(paste0('Gene distance threshold: ', lower_bound, ' <=', thresh))+
      theme_classic()

    return(group.plot)
  })

  grouped_deseq.plot<-wrap_plots(plot.list, ncol = 1)
  ggsave(paste0(figure_filepath, '/motif_role_over_timepoints_rnaseq_between_', x, '.png'), grouped_deseq.plot, height = 15, width = 10)
  ggsave(paste0(figure_filepath, '/motif_role_over_timepoints_rnaseq_between_', x, '.pdf'), grouped_deseq.plot, height = 15, width = 10)
})
```

# Session Information

For the purposes of reproducibility, the R/Bioconductor session information is printed below:

```{r sessioninfo}
sessionInfo()
```












